---
title: "homework2"
format: html
editor: visual
editor_options: 
  chunk_output_type: console
---

## Task 1: Basic Vector Practice

creating vectors pre teatment and post treatment

```{r}
pre_trt_bp <- c(120, 151, 125, 126, 115, 132, 132, 129, 134, 139, 127, 122, 127, 135, 133, 128, 147, 138, 140, 132)

post_trt_bp <- c(127, 145, 135, 122, 115, 122, 123, 126, 126, 129, 132, 146, 120, 114, 121, 120, 128, 120, 133, 115)
```

assign names to the vector elements using past function

```{r}
subjects <- paste("Subject", 1:20, sep = "_")
names(pre_trt_bp) <- names(post_trt_bp) <- subjects 
```

calculate the change in blood pressure for each patient by subtracting post treatment measurements from pretreatment measurements

```{r}
diff_bp <- pre_trt_bp - post_trt_bp
diff_bp
```

calculate the average decrease in blood pressure across all patients using the mean() function

```{r}
mean(diff_bp)
```

determine which patients experiences a decrease in blood pressure after treatment(a positive change). use which() function here

```{r}
diff_bp[diff_bp > 0]
```

calculate the average decrease in blood pressure for those where he blood pressure decreased(positive change)

```{r}
mean(diff_bp[diff_bp > 0])
```

## Task 2: Basic Data Frame Practice

create a data frame object with the four columns corresponding to the data above

```{r}
bp_df <- data.frame(
  patient = subjects, 
  pre_bp = pre_trt_bp, 
  post_bp = post_trt_bp, 
  diff_bp = diff_bp)
bp_df
```

return only rows of data frame where the diff_bp column is negative. use \[\] or learn about subset function. dont reference the original vector from part one acess the column of the data frame

```{r}
bp_df[bp_df$diff_bp < 0, ]
```

Add a new column to the data frame corresponding to TRUE if the post_bp is less than 120 and false if it is not

```{r}
bp_df$normal <- bp_df$post_bp < 120
```

print the data frame nicely to final document

```{r}
knitr::kable(bp_df)
```

## Task 3: List Practice

creating two vectors with pre bp and post bp for placebo group

```{r}
pre_bp_placebo <- c(138, 135, 147, 117, 152, 134, 114, 121, 131, 130)
post_bp_placebo <- c(105, 136, 123, 130, 134, 143, 135, 139, 120, 124)
```

create a new data frame with this data in it that is similar from the data frame in task 2(including new column of true/false)

```{r}
bp_placebo_df <- data.frame(
  patient = subjects[1:10],
  pre_bp = pre_bp_placebo,
  post_bp = post_bp_placebo)
bp_placebo_df$diff_bp <- bp_placebo_df$pre_bp - bp_placebo_df$post_bp
bp_placebo_df$normal <- bp_placebo_df$post_bp < 120
```

now create and store a list with the two elements -- 1st element treatment and 2nd element placebo

```{r}
bp_list <- list(treatment = bp_df, placebo = bp_placebo_df)
bp_list
```

access the first list using 3 different types of syntax

```{r}
bp_list$treatment 
bp_list[1]
bp_list[[1]]
```

in one line of code access the pre_bp column of the placebo data frame

```{r}
bp_list$placebo$pre_bp
```
